---
title: "Untitled"
author: "Thiloshon Nagarajah"
date: "4/21/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ToothGrowth Analysis

Basic Exploratory Analyses 

```{r cars}
library(ggplot2)
f<-ToothGrowth
ggplot(f,aes(supp,len))+facet_grid(.~dose)+geom_boxplot(aes(fill = supp))
```

Just in a glance we can see the dose is influenzing the growth of tooth. The 2 doses is significantly efficient than the 0.5 dosage. But the VC supp shows more efficiency in large doses than in small doses.

Basic summary of data

```{r}
summary(f)
```

Hypothesis Tests

Null Hypothesis 01: There is no difference in tooth growth when different supplements are used.
Alternate Hypothesis of Null Hypothesis 01: There is more tooth growth when using supplement OJ than VC.

```{r}
OJData<- f[f$supp=="OJ",]
VCData<- f[f$supp=="VC",]
t.test(OJData$len , VCData$len, alternative = "greater", paired = FALSE, conf.level = 0.95)
```

We are getting a p-value of 0.03, this is smaller than the rejection p value 0.05. Thus we dont have enough power i.e we dont have enough evidence to reject the alternate hypothesis. This concludes the OJ shows more toothe growth than VC.

Null Hypothesis 02: There is no difference in tooth growth when different dosages are used.
Alternate Hypothesis of Null Hypothesis 02: There is more tooth growth when using dosage 2 than 0.5.


```{r}
OJData<- f[f$dose==0.5,]
VCData<- f[f$dose==2,,]
t.test(VCData$len , OJData$len, alternative = "greater", paired = FALSE, conf.level = 0.95)
```

Here, once again we are getting a p-value of less than 0.05. This shows its highly likely for the alternate hypothesis to be true, which is there is more tooth growth in dose of 2 than 0.5.


Assumptions:

The variables are independent and identically distributed (i.i.d.).
The variables are not from corelated sample
